INLA two dimensional grid replication

### 2022-09-08 src/example_inla-grid/inla-grid.Rmd
Some writing for inla-grid

---
title: "INLA grid"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: true
    toc_collapsed: true
    df_print: paged
    code_folding: show
    theme: lumen
---

```{r}
cbpalette <- multi.utils::cbpalette()
```

Suppose that $\theta = (\theta_1, \theta_2)$ has a multivariate Gaussian distribution with mean vector $\mu$ and covariance matrix $\Sigma$:

@@

ggplot(ground_truth, aes(x = theta1, y = theta2, z = pdf)) +
  geom_contour(col = cbpalette[1]) +
  coord_fixed(xlim = c(-2, 5), ylim = c(-2, 5), ratio = 1) +
  theme_minimal()
```

@@

```{r}
opt <- optim(par = c(0, 0), fn = obj, control = list(fnscale = -1))
opt
```

@@

Create a mapping $\theta(z)$ from the $z$-coordinates to the space of $\theta$:

```{r}
z_to_theta <- function(z) {
  as.vector(opt$par + V %*% sqrt(Lambda) %*% z)
}

theta_mode <- z_to_theta(c(0, 0)) # c(0, 0) maps to the mode
theta_mode

@@

Expand the scaffold in $z$-space:

@@

Map it back to $\theta$-space:

@@

ggplot() +
  geom_point(data = theta_grid, aes(x = theta1, y = theta2), alpha = 0.5) +
  geom_contour(data = ground_truth, aes(x = theta1, y = theta2, z = pdf), col = cbpalette[1]) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend, colour = as.factor(pc)),
               data = segments, size = 1, arrow = arrow(length = unit(0.3, "inches")),
               lineend = "round", linejoin = "round") +
  scale_colour_manual(values = cbpalette[-1]) +
  labs(x = "theta1", y = "theta2") +
  theme_minimal() +
  theme(legend.position = "none")
```

>>>>

---
title: "INLA two dimensional grid replication"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: true
    toc_collapsed: true
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
    **Background** The INLA method uses a grid of hyperparameters to integrate over.

    **Task** We replicate a method for selecting grid vaues as outlined in Rue, Martino and Chopin (2009).
---

## A two dimensional MVN

Suppose that $\theta = (\theta_1, \theta_2)$ has a multivariate Gaussian distribution with mean vector $\mu$ and covariance matrix $\Sigma$ as follows:

@@

ggplot(ground_truth, aes(x = theta1, y = theta2, z = pdf)) +
  geom_contour(col = multi.utils::cbpalette()[1]) +
  coord_fixed(xlim = c(-2, 5), ylim = c(-2, 5), ratio = 1) +
  theme_minimal()
```

@@

```{r}
opt <- optim(par = c(0, 0), fn = obj, control = list(fnscale = -1))
opt$par
```

@@

## Explore the $z$-coordinates

Create a mapping $\theta(z)$ from the $z$-coordinates to $\theta$-coordinates:

```{r}
z_to_theta <- function(z) {
  as.vector(opt$par + V %*% sqrt(Lambda) %*% z)
}
```

The origin `c(0, 0)` maps to the mode:

```{r}
theta_mode <- z_to_theta(c(0, 0))
theta_mode

@@

Expand the scaffold in $z$-coordinates:

@@

Map it back to $\theta$-coordinates:

@@

ggplot() +
  geom_point(data = theta_grid, aes(x = theta1, y = theta2), alpha = 0.5) +
  geom_contour(data = ground_truth, aes(x = theta1, y = theta2, z = pdf), col = multi.utils::cbpalette()[1]) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend, colour = as.factor(pc)),
               data = segments, size = 1, arrow = arrow(length = unit(0.3, "inches")),
               lineend = "round", linejoin = "round") +
  scale_colour_manual(values = multi.utils::cbpalette()[-1]) +
  labs(x = "theta1", y = "theta2") +
  theme_minimal() +
  theme(legend.position = "none")
```
